---
title: "Biostat M280 Homework 4"
subtitle: Due Mar 16 @ 11:59PM
author: Sarah Ji
output: html_document
---

```{r setup, include=FALSE, echo= FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, error = TRUE)
if(!"pacman" %in% row.names(installed.packages())){install.packages("pacman") & load("pacman")}

# these are packages you will need
devtools::install_github("dkahle/ggmap")
pacman::p_load(sparklyr, dplyr, ggplot2, ggmap, maps, mapdata, haven, ggrepel, pacman, lubridate, install = TRUE)
```

```{r}
#1
usa = map_data("usa")
world = map_data("world")
states = map_data("state")
```

```{r}
#2
Sys.setenv(SPARK_HOME="/usr/lib/spark")
config <- spark_config()
sc <- spark_connect(master = "yarn-client", config = config)
```

```{r}
#3
flights_tbl <- tbl(sc, 'flights')
airlines_tbl <- tbl(sc, 'airlines')
airports_tbl <- tbl(sc, 'airports')
```

```{r}
#4
flightzdest = flights_tbl %>% filter(cancelled == 0) %>% select(origin, dest) %>%
group_by(dest) %>% summarise(n_dest = count()) %>%
arrange(desc(n_dest)) %>% collect()

flightzorigin = flights_tbl %>% filter(cancelled==0) %>% select(origin, dest) %>%
group_by(origin) %>% summarise(n_origin = count()) %>%
arrange(desc(n_origin)) %>% collect() 

flightstotal = left_join(flightzorigin, flightzdest, by = c("origin" = "dest")) %>% mutate(N = (n_origin + n_dest)/1000000) %>% arrange(desc(N))

names(flightstotal) = c("Airport", "n_origin", "n_dest", "Total Flights (in millions)")
top10busiestAirports = flightstotal$Airport[1:10]
```

```{r}
#5
top10busiestLatLong = airports_tbl %>% 
filter(faa %in% top10busiestAirports) %>% 
select(AirportID = faa, Airport = name, Latitude = lat, Longitude = lon) %>% collect()

top10busiestLatLong2 = left_join(flightstotal, top10busiestLatLong, by = c("Airport" = "AirportID")) %>% filter(!is.na(Latitude)) %>% mutate(Latitude = as.numeric(Latitude), Longitude = as.numeric(Longitude))
```


Our Apache Yarn cluster hosts the [flights](http://stat-computing.org/dataexpo/2009/the-data.html) data representing 123 million flights over 22 years. Read the [lecture notes](http://hua-zhou.github.io/teaching/biostatm280-2018winter/slides/12-sparklyr/sparklyr-flights.html) on how to access the Yarn cluster. Connect to the database using `sparklyr` and answer following questions. You can base your answers on a specific year or the whole data set.

1. Map the top 10 busiest airports. Size of dots should reflect the number of flights through that destination.  
Hint: You may find this tutorial on [Making Maps in R](http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html) helpful.

```{r}
#6
ggplot() + 
      geom_polygon(data = states, aes(x = long, y = lat, fill = region, group = group), color = "white") + 
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude, size = `Total Flights (in millions)`), color = "black") +
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude, size = `Total Flights (in millions)`), color = "chartreuse") +
      #geom_text(aes(x = Longitude, y = Latitude, label = as.character(top10busiestAirports)), check_overlap = TRUE) +
      scale_size(name = "Total Flights (in millions)", range = c(1, 5)) +
      labs(x = "Longitude", y = "Latitude", title = "Top 10 Busiest Airports in USA") +
      coord_fixed(1.5)
```

**Answer**: The top 10 busiest airports with the most flights going in and out of its facilities are:


2. Map the top 10 busiest direct routes. Size of lines should reflect the number of flights through that route.

```{r, eval = T}
#7
jointpath = flights_tbl %>% filter(cancelled==0 & !is.na(dest) & !is.na(origin)) %>% 
group_by(dest, origin) %>% summarise(n_jointpath = count()) %>%
arrange(desc(n_jointpath)) %>% collect()

top10traveledpathDestination = jointpath$dest[1:10]
top10traveledpathOrigin = jointpath$origin[1:10]

pathlatitudelongitude = airports_tbl %>% 
filter(faa %in% c(top10traveledpathDestination, top10traveledpathOrigin)) %>% 
select(Airport = faa, Latitude = lat, Longitude = lon) %>% collect()

top10buspathwithlatlong1 = left_join(jointpath[1:10, ], pathlatitudelongitude, by = c("dest" = "Airport"))



top10buspathwithlatlong2 = left_join(jointpath[1:10, ], pathlatitudelongitude, by = c("origin" = "Airport"))

top10busiestpathLL = full_join(top10buspathwithlatlong1, top10buspathwithlatlong2, by = "n_jointpath")
testlong = bind_rows(top10buspathwithlatlong1, top10buspathwithlatlong2)


#top10busiestpathLL = top10busiestpathLL[,-c(7:9)]
#names(top10busiestpathLL) = c("Destination", "Origin", "N_path", "Path", "Lat.D", "Long.D", "Lat.O", "Long.O")


long10busypath = testlong %>%
mutate(Latitudes = as.numeric(Latitude), Longitudes = as.numeric(Longitude)) %>% select(Destination = dest, Origin = origin, N_path = n_jointpath, Latitudes, Longitudes)


ggplot() + 
geom_polygon(data = states, 
aes(x = long, y = lat, group = group, fill = region), color = "white") +
coord_fixed(1.3) + 
geom_point(data = long10busypath, aes(x = Longitudes, y = Latitudes), color = "black") + 
geom_curve(data = top10busiestpathLL, aes(x = as.numeric(Longitude.x), y = as.numeric(Latitude.x), xend = as.numeric(Longitude.y), yend = as.numeric(Latitude.y))) +
labs(title = "Map of Top 10 Busiest Routes", 
x = "Longitude (Degrees)", y = "Latitude (Degrees)") 

uniquelabels2 = top10buspathwithlatlong1[!duplicated(top10buspathwithlatlong1[1]),]

#+ geom_label_repel(data = long10busypath,  aes(x = Longitudes, y = Latitudes), label = long10busypath$Destination)
```

**Answer**: The top 10 busiest direct routes with the highest number of flights through that route are:
1) SFO - LAX 
2) LAX - SFO
3) LAX - LAS
4) LAS - LAX
5) PHX - LAX
6) LAX - PHX
7) ORD - MSP
8) MSP - ORD
9) PHX - LAS
10) LAS - PHX



3. LAX:

    (a). Reproduce the plot. Visualize and explain some prominent features you observe. For example, what happened at points 1-5?
    
```{r, eval = T}
#8
LAXflightsbyday = flights_tbl %>% 
  filter(cancelled == 0 & (origin == "LAX" | dest == "LAX") & year >= 1998) %>%
group_by(year, month, dayofmonth) %>% summarise(n = count()) %>% collect()

LAXflightsbyday %>% arrange(year, month, dayofmonth) %>% mutate(date = make_date(year = year, month = month, day = dayofmonth)) %>% ggplot() + geom_line(aes(x = date, y = n)) + coord_cartesian(ylim = c(800, 1400)) + labs(title = "LAX air traffic") +   geom_label(aes(x = make_date(2001, 10, 01), y = 1076, label = "1")) +
      geom_label(aes(x = make_date(2004, 11, 25), y = 1019, label = "2")) +
      geom_label(aes(x = make_date(2004, 07, 04), y = 1060, label = "3")) +
      geom_label(aes(x = make_date(2008, 01, 01), y = 1292, label = "4")) +
      geom_label(aes(x = make_date(2001, 01, 01), y = 1299, label = "5"))

```

**Answer**: We have recreated the desired plot of the number of flights per day from 1998 - 2008, with points labeled 1 to 5 on the dates below:
We notice there is a Holidays effect on the number of flights that come in and out of LAX per day. 9/11 is a historic moment in America,
especially for those working in the airlines industry. While (1) 9/11 is not a holiday that repeats, the rest of the numbered dates are as follows:

Patterns in the data / Borderline Conspiracy Theories:
(1) Fallout of 9/11. A truly tragic piece of American History unfolds and with it comes the fallout..
(2) November 25, 2004, n = 1,019 : Thanksgiving ! Everyone is flying home to see their family members to celebrate thanksgiving.
or ~ The LAX Plan, a law passed on September 29, 2004, to upgrade needed facilities to accomodate current and next generation larger aircraft associated with internationa and long haul domestic travel. https://www.lawa.org/-/media/lawa-web/lawa-our-lax/finallaxplan_092904.ashx
(3) July 4, 2004, n = 1,060 : Independence Day! 
(4) January 1, 2008, n = 1,292 : New Years Day!
or ~ brings on a grim year for a financial crash soon to come, risk averse population.
or ~ The super jumbo jet A380 by Airbus was intended to have its first landing in the United States at the LAX airport in September. 
The A380, is the largest commercial plane that seats 555 passengers, significantly more than it's predecessor, the 747. The airbus could help space-constrained airports, including LAX, by allowing carriers to combine several flights into one.
http://www.latimes.com/travel/la-trw-airbus18mar18-story.html
(5) January 1, 2001, n = 1,299: New Years Day! 


    (b). Visualize and explain seasonal effects.
    
```{r, eval = TRUE}
#9
LAXseason1 = flights_tbl %>% 
  filter(cancelled == 0 & (origin == "LAX" | dest == "LAX") & year >= 1998) %>% group_by(year, month) %>% summarise(peryearmonth = count()) %>% collect()

LAXseason2 = LAXseason1 %>%
      mutate(season = as_factor(labelled(month,
                                      c("Winter" = 12, "Winter" = 1, "Winter" = 2,
                                        "Spring" = 3, "Spring" = 4, "Spring" = 5,
                                        "Summer" = 6, "Summer" = 7, "Summer" = 8,
                                        "Autumn" = 9, "Autumn" = 10, "Autumn" = 11)))) 

LAXseason2 = LAXseason2 %>% group_by(year, season) %>% mutate(n_season = sum(peryearmonth))

    
    LAXseason2 %>% group_by(year) %>% ggplot(aes(x = as.factor(year), y = n_season / 1000000, fill = season)) + 
      geom_bar(stat = "identity") +
      labs(title = "LAX Air Traffic Seasonal Effects",
           x = "Year", y = "Flights (in millions)", fill = "Season") + scale_fill_manual(values = c("lightblue1", "powderblue", "lightblue3", "lightblue4")) + theme_light()

```
  
**Answer**: Above we visualize the number of flights per year (in millions), stratified by season. Notice that the seasonal effect stays consistent over the years.
This is probably the most useful information if you are planning a trip! Summer might be the most optimal time to fly (if you're a student like me), but 
reconsider it when you see how much ticket prices increase for the coveted summer season. 

    (c). Visualize and explain weekly effects.
    
```{r, eval = TRUE}
#10
LAXflightsbyweek = flights_tbl %>% 
  filter(cancelled == 0, origin == "LAX" | dest == "LAX", year >= 1998, !is.na(dayofweek)) %>%
group_by(dayofweek) %>% mutate(n_week = count()) %>% select(year, dayofweek, n_week) %>% collect()

LAXweek1 = flights_tbl %>% 
  filter(cancelled == 0 & (origin == "LAX" | dest == "LAX") & year >= 1998 & !is.na(dayofweek)) %>% group_by(year, dayofweek) %>% 
  summarise(peryearday = count()) %>% collect()

LAXweek2 = LAXweek1 %>%
      mutate(Weekday = as_factor(labelled(dayofweek,
                                      c("Monday" = 1, "Tuesday" = 2, "Wednesday" = 3,
                                        "Thursday" = 4, "Friday" = 5, "Saturday" = 6,
                                        "Sunday" = 7)))) 

LAXweek2 = LAXweek2 %>% group_by(year, Weekday) %>% mutate(n_weekday = sum(peryearday))

    
    LAXweek2 %>% ggplot(aes(x = as.factor(year), y = n_weekday / 1000000, fill = Weekday)) + 
      geom_bar(stat = "identity") +
      labs(title = "LAX Air Traffic Weekly Effects",
           x = "Year", y = "Flights (in millions)", fill = "Weekday") + scale_fill_manual(values = c("#882E72", "#B178A6", "#D6C1DE", "#1965B0", "#5289C7", "#7BAFDE", "#4EB265", "#90C987", "#CAE0AB", "#F7EE55", "#F6C141", "#F1932D", "#E8601C", "#DC050C")) + theme_light()
```
  
**Answer**: Above we visualize the number of flights per year (in millions), by weekday. Notice that the weekly effect stays consistent over the years.

  
    (d). Map top 10 destinations from LAX. Size of dots should reflect the number of flights from LAX to that destination.
    
4. Build a predictive model for the arrival delay (`arrdelay`) of flights flying from LAX. Use the same filtering criteria as in the [lecture notes](http://hua-zhou.github.io/teaching/biostatm280-2018winter/slides/12-sparklyr/sparklyr-flights.html) to construct training and validation sets. You are allowed to use a maximum of 5 predictors. The prediction performance of your model on the validation data set will be an important factor for grading this question.

```{r}
#11
fromLAX3 = flights_tbl %>%
    filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance) & origin == "LAX" & (cancelled == 0)) %>%
    filter(depdelay > 15 & depdelay < 240) %>% #filter outliers
    filter(arrdelay > -60 & arrdelay < 360) %>% #filter outliers
    filter(year >= 2003 & year <= 2007) %>%
    left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>% # merge by the columns
    left_join(airports_tbl, by = c("dest" = "faa")) %>% 
    filter(!is.na(tz_name) & !is.na(alt)) %>%
    select(year, month, arrdelay, depdelay, destination = dest, distance, uniquecarrier, description, tz_name, alt) %>%
group_by(destination) %>% mutate(n_fromLAX = count()) %>% arrange(desc(n_fromLAX))
```

```{r}
#12
# Partition the data into training and validation sets
model_partition <- fromLAX3 %>% 
  sdf_partition(train = 0.8, valid = 0.2, seed = 5555)

# Fit a linear model

ml1 <- model_partition$train %>%
    ml_linear_regression(arrdelay ~ distance + depdelay + uniquecarrier + tz_name + alt)
  
summary(ml1)

```

```{r}
#13
  model_deciles <- lapply(model_partition, function(x) {
    sdf_predict(ml1, x) %>%
      mutate(decile = ntile(desc(prediction), 10)) %>%
      group_by(decile) %>%
      summarize(arrdelay = mean(arrdelay)) %>%
      select(decile, arrdelay) %>%
      collect()
  })
```

```{r}
#14
deciles <- rbind(
  data.frame(data = 'train', model_deciles$train),
  data.frame(data = 'valid', model_deciles$valid),
  make.row.names = FALSE
)
```

```{r}
#15
# Plot average gains by predicted decile
deciles %>%
  ggplot(aes(factor(decile), arrdelay, fill = data)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Average Arrival Delay by Predicted Decile', x = 'Decile', y = 'Minutes')
```
    
```{r}
#16
data_2008 = flights_tbl %>%
    filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance) & origin == "LAX" & (cancelled == 0)) %>% filter(depdelay > 15 & depdelay < 240) %>% #filter outliers
    filter(arrdelay > -60 & arrdelay < 360) %>% #filter outliers
    filter(year == 2008) %>%
    left_join(airports_tbl, by = c("dest" = "faa")) %>% 
    filter(!is.na(tz_name) & !is.na(alt)) %>%
    select(year, month, arrdelay, depdelay, dest, distance, uniquecarrier, tz_name, alt)


decile2008 <- sdf_predict(x = data_2008, model = ml1) %>% select(arrdelay, uniquecarrier, prediction)

```


5. Visualize and explain any other information you want to explore.

```{r}
#17
pre911 = model_data_pre911 %>% group_by(description , year) %>% summarise(arrivaldelay = mean(arrdelay))

post911 = model_data %>% group_by(description , year) %>% summarise(arrivaldelay = mean(arrdelay))

pre911 %>% group_by(year, description) %>% ggplot(aes(x = as.factor(year), y = mean(arrivaldelay))) + geom_bar(stat = "identity") +
      labs(title = "LAX Air Traffic Pre - 9/11",
           x = "Year", y = "Mean Arrival Delay")

post911 %>% group_by(year, description) %>% ggplot(aes(x = as.factor(year), y = mean(arrivaldelay))) + geom_bar(stat = "identity") +
      labs(title = "LAX Air Traffic Post - 9/11",
           x = "Year", y = "Mean Arrival Delay")

model_data %>%
  mutate(id = row_number()) %>%
  select(x, y, z, id) %>%
  gather(variable, value, -id)  %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 500) +
  geom_rug() +
  facet_grid(variable ~ .)

```

```{r}
#deciles %>% mutate(sum = sum()) %>%
#    ggplot(aes(sched_dep_time)) +
#    geom_density() +
#    facet_grid(~cancelled) + 
#    labs(title = "Departure Time Cancelled (T) vs. Not Cancelled (F) Flights")

data.frame(data = 'train', data_2008)

ggplot() + geom_boxploth(aes(x = tz_name, y = arrdelay,
                                     colour = cut_number(distance , 10)))

LAXseason2 %>% group_by(year) %>% ggplot(aes(x = as.factor(year), y = n_season / 1000000, fill = season)) + geom_bar(stat = "identity") +
      labs(title = "LAX Air Traffic Seasonal Effects",
           x = "Year", y = "Flights (in millions)", fill = "Season") + scale_fill_manual(values = c("lightblue1", "powderblue", "lightblue3", "lightblue4")) + theme_light()

```


**Answer**:  


```{r, evaluate = T, include = FALSE, echo = FALSE}
spark_disconnect(sc)
```