---
title: "hw4_280"
author: "Sarah Ji"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
#http://35.185.198.142:8787/
knitr::opts_chunk$set(echo = TRUE)
library(sparklyr)
library(dplyr)
library(ggplot2)
# these are packages you will need, but probably already have.
# Don't bother installing if you already have them
#install.packages(c("ggplot2", "devtools", "dplyr", "stringr"))

# some standard map packages.
#install.packages(c("maps", "mapdata"))

# the github version of ggmap, which recently pulled in a small fix I had
# for a bug 
devtools::install_github("dkahle/ggmap")
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
```

```{r}
usa = map_data("usa")
world = map_data("world")
```


```{r}
Sys.setenv(SPARK_HOME="/usr/lib/spark")
config <- spark_config()
sc <- spark_connect(master = "local", config = config)
```

Access hive tables
```{r}
# Cache flights Hive table into Spark
#tbl_cache(sc, 'flights')
flights_tbl <- tbl(sc, 'flights')

#show what columns I have lazily
#flights_tbl %>% print(width = Inf)
```

```{r}
# Cache airlines Hive table into Spark
#tbl_cache(sc, 'airlines')
airlines_tbl <- tbl(sc, 'airlines')
#airlines_tbl %>% print(width = Inf)
```

```{r}
# Cache airports Hive table into Spark
#tbl_cache(sc, 'airports')
airports_tbl <- tbl(sc, 'airports')
#airports_tbl %>% print(width = Inf)
```

Map the top 10 busiest airports. Size of dots should reflect the number of flights through that destination.
Hint: You may find this tutorial on Making Maps in R helpful.

```{r}
flightzz <- flights_tbl %>% select(year, uniquecarrier, flightnum, origin, dest, cancelled) %>% collect()

flightzz <- flights_tbl %>% select(year, arrdelay, depdelay, distance, uniquecarrier, flightnum, origin, dest, cancelled) %>% collect()

flightzz = flights_tbl %>% filter(cancelled == 0) %>% select(origin, dest, cancelled) %>%
group_by(dest) %>% summarise(n_dest = count()) %>%
arrange(desc(n_dest)) %>% collect()

flightzorigin = flights_tbl %>% filter(cancelled==0) %>% select(origin, dest, cancelled) %>%
group_by(origin) %>% summarise(n_origin = count()) %>%
arrange(desc(n_origin)) %>% collect()

flightstotal = left_join(flightzorigin, flightzz, by = c("origin" = "dest")) %>% mutate(N = (n_origin + n_dest)/1000000) %>% arrange(desc(N))

names(flightstotal) = c("Airport", "n_origin", "n_dest", "Total Flights (in millions)")
top10busiestAirports = flightstotal$Airport[1:10]

flights_cleanOrigin <- flights_tbl %>%
filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance) & (cancelled == 0)) %>%
  filter(depdelay > 15 & depdelay < 240) %>%
  filter(arrdelay > -60 & arrdelay < 360) %>% 
  group_by(origin) %>% summarise(n_origin = count()) %>%
arrange(desc(n_origin)) %>% collect()

flights_clean <- flights_tbl %>%
filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance) & (cancelled == 0)) %>%
  filter(depdelay > 15 & depdelay < 240) %>%
  filter(arrdelay > -60 & arrdelay < 360) %>% 
  group_by(dest) %>% summarise(n_dest = count()) %>%
arrange(desc(n_dest)) %>% collect()
  
```

```{r}
ggplot() + geom_polygon(data = usa, aes(x = long, y = lat, group = group), fill = "lavender", color = "blue") + 
coord_fixed(1.3) 
```

```{r}
top10busiestLatLong = airportz %>% 
filter(faa %in% top10busiestAirports) %>% 
select(AirportID = faa, Airport = name, Latitude = lat, Longitude = lon) %>% collect()

top10busiestLatLong2 = left_join(flightstotal, top10busiestLatLong, by = c("Airport" = "AirportID")) %>% filter(!is.na(Latitude)) %>% mutate(Latitude = as.numeric(Latitude), Longitude = as.numeric(Longitude))
```

```{r}
ggplot() + 
      geom_polygon(data = usa, aes(x = long, y = lat), fill = "lavender", color = "blue") + 
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude), color = "black", size = 5) +
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude), color = "chartreuse", size = 4) +
      coord_fixed(1.3)
```


```{r}
states = map_data("state")
ggplot() + 
      geom_polygon(data = states, aes(x = long, y = lat, fill = region, group = group), color = "white") + 
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude, size = `Total Flights (in millions)`), color = "black") +
      geom_point(data = top10busiestLatLong2, aes(x = Longitude, y = Latitude, size = `Total Flights (in millions)`), color = "chartreuse") +
      #geom_text(aes(x = Longitude, y = Latitude, label = as.character(top10busiestAirports)), check_overlap = TRUE) +
      scale_size(name = "Total Flights (in millions)", range = c(1, 5)) +
      labs(x = "Longitude", y = "Latitude", title = "Top 10 Busiest Airports in USA") +
      coord_fixed(1.5)

```


```{r}

```








```{r}
airlinez <- airlines_tbl %>% collect()
```

```{r}
airportz <- airports_tbl %>% group_by(faa) %>% collect()
```


Data starts around 1987 first year may not have complete data. 22 years of data. 

```{r}
out
```

Barplot of the number of data points in each year

```{r}
out %>% ggplot(aes(x = year, y = n)) + geom_col()
```



```{r}
origin_airports <- flights_tbl %>%
  group_by(origin) %>%
  mutate(n1 = n()) %>% collect()

destination_airports <- flights_tbl %>%
  group_by(dest) %>%
  mutate(n2 = n()) %>% collect()


destination_airports %>% left_join(origin_airports, by = "flightnum") %>% mutate(n = n1 + n2)
```

```{r}
out
```

Number of flights taking off from LAX:
question: what is the jump 1996 and dip in 2008
```{r}
out %>% ggplot(aes(x = year, y = n)) + 
    geom_col() +
    labs(title = "Number of flights from LAX")
```

How many flights to LAX per year:

```{r}
system.time({
out <- flights_tbl %>%
  filter(dest == "LAX") %>%
  group_by(year) %>%
  count() %>%
  arrange(year) %>%
  collect()
})
```

FIT A MODEL:

Suppose we want to fit a linear regression of gain (depdelay - arrdelay) on distance, departure delay, and carriers using data from 2003-2007.

not collected because it will be huge!!
```{r}
# Filter records and create target variable 'gain'
system.time(
  model_data <- flights_tbl %>%
    filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
    filter(depdelay > 15 & depdelay < 240) %>% #filter outliers
    filter(arrdelay > -60 & arrdelay < 360) %>% #filter outliers
    filter(year >= 2003 & year <= 2007) %>%
    left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>% # merge by the columns
    mutate(gain = depdelay - arrdelay) %>%
    select(year, month, arrdelay, depdelay, distance, uniquecarrier, description, gain)
)
```

```{r}
model_data
```

Get descriptive statistics: for every airline summarize gain distance depdelay

```{r}
# Summarize data by carrier
model_data %>%
  group_by(uniquecarrier) %>%
  summarize(description = min(description), gain = mean(gain), 
            distance = mean(distance), depdelay = mean(depdelay)) %>%
  select(description, gain, distance, depdelay) %>%
  arrange(gain)
```

**FIT LINEAR REGRESSION**
Train a linear model
Predict time gained or lost in flight as a function of distance, departure delay, and airline carrier.

split to training and testing data

```{r}
# Partition the data into training and validation sets
model_partition <- model_data %>% 
  sdf_partition(train = 0.8, valid = 0.2, seed = 5555)

# Fit a linear model on training data with gain as response
system.time(
  ml1 <- model_partition$train %>%
    ml_linear_regression(gain ~ distance + depdelay + uniquecarrier)
)

```

```{r}
# Summarize the linear model
summary(ml1)
```

Assess model performance
Compare the model performance using the validation data.

Apply the model partition separatately on train and test data. 

```{r}
# Calculate average gains by predicted decile
system.time(
  model_deciles <- lapply(model_partition, function(x) {
    sdf_predict(ml1, x) %>%
      mutate(decile = ntile(desc(prediction), 10)) %>%
      group_by(decile) %>%
      summarize(gain = mean(gain)) %>%
      select(decile, gain) %>%
      collect()
  })
)
```

```{r}
model_deciles
```

```{r}
# Create a summary dataset for plotting
deciles <- rbind(
  data.frame(data = 'train', model_deciles$train),
  data.frame(data = 'valid', model_deciles$valid),
  make.row.names = FALSE
)
deciles
```

visualize performance of the model it looks good.

```{r}
# Plot average gains by predicted decile
deciles %>%
  ggplot(aes(factor(decile), gain, fill = data)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Average gain by predicted decile', x = 'Decile', y = 'Minutes')
```

Visualize predictions
Compare actual gains to predicted gains for an out of time sample.

new dataset
```{r}
# Select data from an out of time sample
data_2008 <- flights_tbl %>%
  filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
  filter(depdelay > 15 & depdelay < 240) %>%
  filter(arrdelay > -60 & arrdelay < 360) %>%
  filter(year == 2008) %>%
  left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>%
  mutate(gain = depdelay - arrdelay) %>%
  select(year, month, arrdelay, depdelay, distance, uniquecarrier, description, gain, origin, dest)
data_2008
```

for each airline plot the predicted vs actual values of gain. southwest and united wait.
delta is good. 

```{r}
# Plot actual gains and predicted gains by airline carrier
ggplot(carrier, aes(gain, prediction)) + 
  geom_point(alpha = 0.75, color = 'red', shape = 3) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
  geom_text(aes(label = substr(description, 1, 20)), size = 3, alpha = 0.75, vjust = -1) +
  labs(title='Average Gains Forecast', x = 'Actual', y = 'Predicted')
```

```{r}
spark_disconnect_all()
```

